Canonical rings of $\bq$-divisors on Minimal Rational Surfaces

### 2015-07-17 surfaces-write-up/surface-write-up.tex
Changes to 5.3

To find such a $\beta$, we simply write $f_i$ as a 
polynomial in $f_1, \ldots f_4$. Furthermore, these
generate all relations, since they allow us to reduce any $\prod_{j =
1}^n f_j^{r_j}$ to the canonical form where $r_i = - \lfloor d \alpha_i \rfloor$ whenever $i > 4$.

%Note that we can view a relation from Equation
%~\ref{eqn:hirz-relations-psi} as an element of $S$ given
%by $(d, c_1, \ldots, c_{i-1}, c_i + 1, c_{i+1}, \ldots , c_n)$.  
Let $S'\subseteq S$ be the
set of $(d, c_1, \ldots, c_{i-1}, c_i + 1, c_{i+1}, \ldots , c_n)$ satisfying $c_i \ge -\alpha_i d$,
$\bida_i + \sum_{j = 1}^m \bida_j c_j = 0$, and $\bidb_i + \sum_{j=1}^n \bidb_j c_j
= 0$.  A careful computation deduces that the lattice point corresponding to $y_i \prod_{j=1}^n {y_j}^{c_j}$ is of the form 
\[
	E + \sum_{k\in \Te} r_k e_k + \sum_{\substack{h \in \Tp \\ k \in \Tm}} r_{h,k} e_{h,k}
\]

\noindent
where $E$ is in degree $\mathbb{R}_{\ge 0}$
combination of elements in 
\[
	\{e_j: j\in \{i\} \cap S\} \cup \{e_{i, j}: j \in \Tm \mbox{ if }
	i	\in \Tp \} \cup \{e_{j, i}: j \in \Tp \mbox{ if } i \in \Tm \}.
\]

Let $M$ be the set of monomials of $\beta_j$.  Then for $\mu\in M$, the corresponding lattice point is of the form
\[
	E_\mu + \sum_{k\in \Te} r_k e_k + \sum_{\substack{
	h \in \Tp \\ k \in \Tm}} r_{h,k} e_{h,k}
\]
where $E_\mu$ is a $\mathbb{R}_{\ge 0}$-combination of elements in
\begin{align*}
	\{e_j : j \in \{1, 2, 3, 4\} \cap \Te\}
	&\cup \; \{e_{j, k}: k \in \Tm, \; j \in \{1, 2, 3, 4\} \cap \Te\} \\
	&\cup \; \{e_{k, j}: k \in \Tp, \; j \in \{1, 2, 3, 4\} \cap \Tm\}.
\end{align*}

>>>>

To find such a $\beta$, we simply write $f_i$ as a 
polynomial in $f_1, \ldots f_4$. Furthermore, these
generate all relations, since they allow us to reduce any $\prod_{j =
1}^n f_j^{c_j}$ to the canonical form where $r_i = - \lfloor d \alpha_i \rfloor$ whenever $i > 4$. \todo{Add $\ell_i$ defenition}

%Note that we can view a relation from Equation
%~\ref{eqn:hirz-relations-psi} as an element of $S$ given
%by $(d, c_1, \ldots, c_{i-1}, c_i + 1, c_{i+1}, \ldots , c_n)$.  
Let $\Sigma'\subseteq \Sigma$ be the \todo{change all $S$ to sigma's}
set of $(d, c_1, \ldots, c_{i-1}, c_i + 1, c_{i+1}, \ldots , c_n)$ satisfying $c_i \ge -\alpha_i d$,
$\bida_i + \sum_{j = 1}^m \bida_j c_j = 0$, and $\bidb_i + \sum_{j=1}^n \bidb_j c_j
= 0$.  Let $\sigma\in \Sigma'$ correspond to the monomial $y_i \prod_{j=1}^n {y_j}^{c_j}$.  Then we can write $\sigma$ as a sum of $r_j e_j$'s for $j\in \Te$ and $r_{j,k}$ for $r\in \Tp, k\in \Tm$.  Let $V$ be the set of $e_i$ $e_{i,j}$ and $e_{j,i}$ as $j$ ranges that exist (i.e. $e_i$ if the hyperplane from lemma \ref{lem:hirz-generators} intersects $e_i$, and the $e_{i,j}$'s or $e_{j,i}$'s if the hyperplane intersects the corresponding edge of the simplex).  A careful computation deduces that if $e_i$ exists
\[
	l
\]
\[
	\lambda + E
\]

\noindent
where $\lambda\in S$ and
\[
	k
\]

\[
	\deg(E) \le \begin{cases}
	\ell_i \gcd(\bida_i, \bidb_i)	&\mbox{ if } i \in \Te \\
	\max_{j \in \Tm} \bigl(\ell_{i, j} (\bida_i \bidb_j - \bida_j \bidb_i)\bigr)
	&\mbox{ if } i \in \Tp \\
	\max_{j \in \Tp} \bigl(\ell_{j, i} (\bida_j \bidb_ i - \bida_i \bidb_j) \bigr)
	&\mbox{ if } i \in \Tm \end{cases}.
\]

 $= V_1 \cup V_2 \cup V_3$ where
\begin{align*}
	&	V_1 = \{ \{e_i\} \mbox{ if }
	&	V_2= \{e_{i, j}: j \in \Tm \mbox{ if } i	\in \Tp \}
	&	V_3=  \{e_{j, i}: j \in \Tp \mbox{ if } i \in \Tm \}.
\end{align*}

Let $M$ be the set of monomials of $\beta_j$.  Then for $\mu\in M$, the corresponding lattice point is of the form
\[
	E_\mu + \sum_{k\in \Te} r_k e_k + \sum_{\substack{
	h \in \Tp \\ k \in \Tm}} r_{h,k} e_{h,k}
\]
where $E_\mu$ is in degree
\[
	\deg(F) \le \begin{cases}
	\ell_i \gcd(\bida_i, \bidb_i)	&\mbox{ if } i \in \Te \\
	\max_{j \in \Tm} \bigl(\ell_{i, j} (\bida_i \bidb_j - \bida_j \bidb_i)\bigr)
	&\mbox{ if } i \in \Tp \\
	\max_{j \in \Tp} \bigl(\ell_{j, i} (\bida_j \bidb_ i - \bida_i \bidb_j) \bigr)
	&\mbox{ if } i \in \Tm \end{cases}.
\]

a $\mathbb{R}_{\ge 0}$-combination of elements in
\begin{align*}
	\{e_j : j \in \{1, 2, 3, 4\} \cap \Te\}
	&\cup \; \{e_{j, k}: k \in \Tm, \; j \in \{1, 2, 3, 4\} \cap \Te\} \\
	&\cup \; \{e_{k, j}: k \in \Tp, \; j \in \{1, 2, 3, 4\} \cap \Tm\}.
\end{align*}
